A Two-Infall Star Formation History with Outflows and Radial Migration

### 2024-12-12 src/tex/ms.tex

@@

@@

\citet{chiappini_chemical_1997} presented the first two-infall model of the MW disk. They adopted an inside-out prescription for the formation of the thin disk, with the timescale of the second infall increasing linearly with radius. They assume the formation timescale of the thick disk is 1 Gyr, and the time of the second infall is 2 Gyr. They also assume a lower threshold on the gas density required for star formation, which produces a star formation hiatus at the end of the thick disk phase. They also noted the ``loop'' that appears in the \aFe--[Fe/H] relation during the thin disk phase.

@@

% Radial gas flows
Some two-infall studies \citep[e.g.,][]{palla_chemical_2020,palla_mapping_2024} implement inward radial gas flows with velocity $\sim1$ km s$^{-1}$ in order to reproduce the radial abundance gradient without Galactic outflows.

@@

\section{One-Zone Models}
\label{sec:onezone-results}

@@

>>>>

\section{Outline}

\subsection{Outline of Key Plots}

\begin{itemize}
    \item The turn-over in [O/Fe] evolution due to the second infall epoch produces a peak in the stellar distribution at intermediate [O/Fe], which is not observed in APOGEE.
    \item Attempts to fix the problem with the [O/Fe] distribution in the thin disk end up creating issues with other observables, such as the thick/thin disk surface density ratio, the SN Ia DTD, or the age-metallicity relation.
    \item The equilibrium model, if accurate, places very strict limits on the two-infall model, moreso than for other models.
    \item A short ($\sim200$ Myr) star formation hiatus early in the history of the disk is sufficient to produce a bimodal distribution of [O/Fe] without altering the gas infall prescription.
\end{itemize}

@@

% To add: Karma Gallart (2019) three-infall-ish, Kobayashi metallicity limit for SNe Ia and sub-Chandra

@@

\citet{chiappini_chemical_1997} presented the first two-infall model of the MW disk. They adopted an inside-out prescription for the formation of the thin disk, with the timescale of the second infall increasing linearly with radius. They assume the formation timescale of the thick disk is 1 Gyr, and the time of the second infall is 2 Gyr. They also assume a lower threshold on the gas density required for star formation, which produces a star formation hiatus at the end of the thick disk phase. They also noted the ``loop'' that appears in the \aFe--[Fe/H] relation during the thin disk phase. They mostly compare to Solar neighborhood data.

\citet{chiappini_abundance_2001} updated their model to fit the observed abundance, stellar, and gas gradients in the Milky Way disk. They find that a density threshold for star formation, making a gap in star formation between the halo and disk phases, is necessary to reproduce observational data. They argue that the details of halo formation also have a strong impact on the outer disk. No radial flows or outflows. I'm confused about their distinction between halo and thick disk, if any.

\citet{chiappini_oxygen_2003} compare model predictions of C, N, and O to Milky Way and other spiral galaxy data. They predict a very slow enrichment between the birth of the Sun and present day, which they claim is due to their adopted gas density threshold for star formation (the infall rate is low enough that the SFR oscillates from timestep to timestep).

\citet{matteucci_effect_2009} investigated different SN Ia delay time distributions using the \citep{chiappini_chemical_1997,chiappini_abundance_2001} model. They use a model for the full disk but compare only to Solar neighborhood data. They concluded that some prompt SN Ia enrichment is necessary to match observed abundance patterns.

\citet{spitoni_effects_2009} relaxed the instantaneous recycling approximation in two different ways: galactic fountains (i.e., SN ejecta are launched out of the disk and take some time to return) and metal cooling (i.e., newly-enriched hot gas takes time to cool before it can form stars). They found that fountains do not strongly affect the evolution of the $\alpha$-elements for short (few hundred Myr) timescales. The effect of a metal cooling timescale is similarly negligible if the yields are not dependent on metallicity.

\citet{spitoni_effects_2011} find that a two-infall model of the disk without gas exchange produces a radial metallicity gradient which is too shallow. They implement an inward radial gas flow on the order of $\sim0-4$ km s$^{-1}$ which varies with radius, and find that it improves agreement with the observed gradient. However, they found that a variable star formation efficiency with radius in combination with a gas density threshold for star formation could also reproduce the observed gradient without radial flows.

\citet{spitoni_effect_2015}

@@

% Radial gas flows
Some two-infall studies \citep[e.g.,][]{spitoni_effects_2011,palla_chemical_2020,palla_mapping_2024} implement inward radial gas flows with velocity $\sim1$ km s$^{-1}$ in order to reproduce the radial abundance gradient without Galactic outflows.

@@

\section{Exploring the Parameter Space in One-Zone Models}
\label{sec:onezone-results}

@@

Figure \ref{fig:onezone-dtd} illustrates the effect of the SN Ia DTD on the abundance predictions by the two-infall model. The fiducial plateau DTD produces the strongest high-$\alpha$ peak as well as the greatest separation between the two low-$\alpha$ peaks. The exponential DTD produces the most stars near Solar [O/Fe], but the low-$\alpha$ sequence is still distinctly bimodal. Finally, the power-law DTD shrinks the low-$\alpha$ loop such that there is only one low-$\alpha$ peak; however, this DTD produces high-$\alpha$ stars at metallicities much lower than observed, as discussed in \citet{dubay_galactic_2024}.

\begin{figure*}
    \centering
    \includegraphics{figures/onezone_dtd.pdf}
    \caption{The model parameters are fixed at $\tau_1=1$ Gyr, $\tau_2=15$ Gyr, and $\tau_{\rm on}=3.2$ Gyr.}
    \label{fig:onezone-dtd}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics{figures/yield_outflow.pdf}
    \caption{Caption.}
    \label{fig:onezone-dtd}
\end{figure}
